Make maps of loan words

### 2025-03-27 analysis/paper/paper.qmd
add archaeological points

@@

  # Get the background map using maptiles package
  background_map <- maptiles::get_tiles(data_sf, provider = "Esri.WorldGrayCanvas", crop = TRUE)

  # Increase the boundaries of the plot
  bbox <- sf::st_bbox(data_sf)
  x_range <- bbox["xmax"] - bbox["xmin"]
  y_range <- bbox["ymax"] - bbox["ymin"]
  bbox["xmin"] <- bbox["xmin"] - 50 * x_range
  bbox["xmax"] <- bbox["xmax"] + 50 * x_range
  bbox["ymin"] <- bbox["ymin"] - 50 * y_range
  bbox["ymax"] <- bbox["ymax"] + 50 * y_range

  # Create the tmap object
  tmap::tmap_mode("plot")
  map_plot <- tmap::tm_shape(background_map) +
    tmap::tm_rgb() +
    tmap::tm_shape(data_sf) +
    tmap::tm_text(text = form_column, col = style_column, palette = dark_palette) +
    tmap::tm_layout(title = plot_title, legend.show = FALSE, bbox = bbox)

>>>>

### Add further layers to the map
Get the archaeological data layers
```{r}
# Create a vector of file paths
points <- here::here("analysis", "data", "raw_data", c("gold_points.csv", 
                                                     "fonio_points.csv",
                                                     "millet_points.csv",
                                                     "salt_points.csv",
                                                     "iron_points.csv"))

# read the csv files into a list of data frames
point_list <- purrr::map(points, readr::read_csv2)

# make sf geometry
test <- point_list[[1]] |> sf::st_as_sf(coords = c("lon","lat"))

# set crs
sf::st_crs(test) <- 4326
```

function to add further layers
```{r}
add_sf_to_leaflet_map <- function(leaflet_map, new_data_sf, label_column) {

  # Ensure the new input data is an sf object and contains the required columns
  if (!inherits(new_data_sf, "sf")) {
    stop("The new input data must be an sf object.")
  }
  if (!label_column %in% colnames(new_data_sf)) {
    stop(paste("The column", label_column, "is not found in the new data."))
  }

  # Define a color for the new points
  new_point_color <- "black"

  # Add the new sf object as a point layer to the existing leaflet map
  leaflet_map <- leaflet_map |>
    leaflet::addCircleMarkers(
      data = new_data_sf,
      lng = ~sf::st_coordinates(geometry)[, 1],
      lat = ~sf::st_coordinates(geometry)[, 2],
      color = new_point_color,
      radius = 2,
      stroke = FALSE, fillOpacity = 1,
      label = purrr::map(
        glue::glue("<span style='color:{new_point_color}; font-weight: normal; font-size: 10px;'>{as.character(new_data_sf[[label_column]])}</span>"),
        htmltools::HTML
      ),
      labelOptions = leaflet::labelOptions(
        noHide = TRUE,
        direction = 'top',
        textOnly = TRUE
      )
    )

  # Return the updated map object
  return(leaflet_map)
}
```

```{r}
#run mapping function
create_leaflet_map(data_list$gold, background = "bw") |>

#add point layer
add_sf_to_leaflet_map(test, label_column = "site")
```

@@

  # Increase the boundaries of the plot
  bbox <- sf::st_bbox(data_sf)
  x_range <- bbox["xmax"] - bbox["xmin"]
  y_range <- bbox["ymax"] - bbox["ymin"]
  bbox["xmin"] <- bbox["xmin"] - 0.5 * x_range
  bbox["xmax"] <- bbox["xmax"] + 0.5 * x_range
  bbox["ymin"] <- bbox["ymin"] - 0.5 * y_range
  bbox["ymax"] <- bbox["ymax"] + 0.5 * y_range
  attr(bbox, "class") <- "bbox"

  # Get the background map using maptiles package without cropping
  background_map <- maptiles::get_tiles(data_sf, provider = "Esri.WorldGrayCanvas", crop = FALSE)

  # Create the tmap object
  tmap::tmap_mode("plot")
  map_plot <- tmap::tm_shape(background_map, bbox = bbox) +
    tmap::tm_rgb() +
    tmap::tm_shape(data_sf) +
    tmap::tm_text(text = form_column, col = style_column, 
                  col.scale = tmap::tm_scale(values = dark_palette)) +
    tmap::tm_layout(title = plot_title, legend.show = FALSE, bbox = bbox)
